It's Never Too Late To Learn\\ \emph{Using Online Machine Learning in Distributed Systems}

### 2016-01-22 writeups/msr_workshop/paper.tex
Modifications, mainly to section 1

Building a modern distributed service that guarantees service-level objectives
is hard. Complexity is introduced by the deployment environment and exacerbated
as the service evolves with time to accommodate hardware/environment changes or
support new features.
For example, to build a simple caching layer that guarantees a certain hit rate,
response latency, and throughput, developers need to take into account the expected
workload hitting the cache, the number of machines available to run it, the network
connecting these machines, the storage hierarchy within each machine, and interference
from other applications running on the network and machines.  A change in any one
of these factors could render the service inefficient.

Traditionally, this complexity is addressed by using highly-skilled developers
to build custom systems that are optimized for specific configurations and
objective functions.  Is there a general, principled approach for dealing with
complexity in distributed systems? Hand-tuning the design and configuration of a system
does not scale well with increased complexity and changes in software design, hardware
architecture, deployment environment, and service-level objectives.

We propose to use online learning to cope with this complexity. Specifically, we
view a {\em decision} made by a distributed system in the framework of
contextual online learning with partial feedback. This is a natural fit: we typically
have plenty of context surrounding a decision (which we may not know how to use), the decision
is made online, and we only receive feedback for the decision that was made. Thus, we can
replace a hand-designed policy with a new policy that has been optimized by a contextual
learning system. By supplying rich contexts to this policy, we can capture the complexity of
our environment without having to understand it; by optimizing the policy continuously, we can
can adapt to changes in the environment or service over time.

Machine learning has certainly been used in the past to optimize system decisions. Most
of this use has been restricted to offline settings with full information: for example,
supervised learning is often used to train models on data that has been annotated with
correct labels. Some systems work in an online setting with full information: for example,
a cache that applies an experts learning framework to dynamically switch between existing
(hand-designed) cache eviction policies, by running all policies in parallel and measuring
their performance. Systems that work in an online setting with partial information tend to
be for news or ad recommendation, but examples from systems infrastructure exist: for example,
caching dynamic query content [], tuning parameters of multicore data structures [], allocating
resources to cluster computing tasks [], and allocating servers to web applications [].
In all of these systems, the context used to make a decision, and the
decision itself, are {\em local} to a machine. This is achieved by either devising a decentralized
algorithm in the first place, or giving up on global optimality. Thus we are not aware of any
system that makes ``global'' decisions based on truly distributed architecture or state.

And yet we use hand-designed policies to make these decisions all the time, such as for request
routing, replica placement, replica selection, cloud resource allocation, failure recovery tasks, 
and others. What is it about distributed systems that makes it difficult to apply online learning
at standard decision points? This proposal identifies fundamental challenges posed by distributed
systems that must be analyzed and addressed in order to make online learning a standard tool for
cloud optimization.  These challenges hold even if we assume that the online learning system is
a black-box that delivers optimal decisions with zero latency--in other words,
they are fundamental to the way we design distributed systems.  We begin by highlighting some of the unknown factors
that contribute to these challenges, and then list the challenges themselves.

@@

Systems often make decisions without complete knowledge of the factors that affect,
or could affect, those decisions.  For example, a caching system evicts items based
on what it believes future requests will look like, without actually knowing the future.
In other cases, incomplete knowledge is due to the absence of data-computation locality.
For example, a load balancer makes decisions based on what it thinks the current load on
machines is, e.g. based on data it collected recently, rather than the actual load at that
very instant in time.

In both of these situations, assumptions are made about the unknowns offline during
\emph{design time}, resulting in different routing and caching algorithms. A change to
the \emph{run-time} environment could easily render these choices suboptimal.

@@

In some situations, system designers have no way of knowing deployment details
that might impact a system's performance. For example, a system could be
built for external deployment and used by third-party customers, or it could be
deployed in a multi-tenant environment where co-located services could
affect its performance in unexpected ways.

In these cases, it is nearly impossible for developers to account for all 
unknown variables at design time.

@@

As mentioned earlier, we adopt the framework of contextual online learning
with partial feedback.  To separate the systems concerns we are trying to
highlight from machine learning concerns, we abstract the learning system as a
black-box and assume it provides optimal decisions with zero latency. For a
given decision over a discrete action space, the interface to the learning
system is as follows: given a {\em context} that summarizes the state of the
world, the learning system chooses an {\em action} to take, which our system
does and later reports a {\em reward} indicating how good the action was. An
example of a learning system that supports this interface is the Decision
Service~\footnote{http://aka.ms/mwt}. An example of a distributed systems
decision is routing requests at a load balancer: the context could be load
information of each machine in the cluster, the action could be one of the
possible machines, and the reward could be the latency of processin the request.

Thus we have reduced our concerns to generating an appropriate context,
executing an action (told to us by the learnin system), and reporting reward
information.  We believe that fundamental properties of distributed systems make
even these tasks difficult, as described below.

@@

\subsection*{Hierarchy}
Distributed systems are often designed in a hierarchical manner, where
subcomponents are monitoring and managing others. These components and layers of
indirection simplify reasoning about the system at large. However, it is not
clear how hierarchy will affect an ML agent.
\begin{itemize}
  \item Will hierarchy aid the learning system via parallelism? Or will it hinder
    it since not all features are visible at every level?
  \item Should learning happen in each component of the system? Or should it be
    centralized?
  \item If learning is happening at multiple levels of the hierarchy, will that
    create interference between the learners? Or will they complement each
    other?
\end{itemize}

- Should feature vectors cut across the levels or not?
- In general should be composable if at each level can define decision and reward,
and exploration randomization is different at each level

>>>>

Building a modern distributed service that guarantees service-level objectives
is hard. Complexity is introduced by the deployment environment and exacerbated
as the service evolves with time to accommodate hardware/environment changes or
support new features.
For example, to build a simple caching layer that guarantees a certain hit rate,
response latency, and throughput, developers need to take into account the
expected workload accessing the cache, the number of machines available to run
it, the network connecting these machines, the storage hierarchy within each
machine, and interference from other applications running on the network and
machines.  A change in any one of these factors could render the service
woefully inefficient.

Traditionally, this complexity is addressed by using highly-skilled developers
to build custom systems that are optimized for specific configurations and
objective functions.  Is there a general, principled, approach for dealing with
complexity in distributed systems? Hand-tuning the design and configuration of a
system does not scale well with increased complexity or changes in software
design, hardware architecture, deployment environment, and service-level
objectives.

We propose to use online learning to cope with this complexity. Specifically, we
view a {\em decision} made by a distributed system in the framework of
contextual online learning with partial feedback. This is a natural fit for
distributed systems where: there is plenty of context surrounding each decision,
decisions are made online, and feedback is received only for decisions that were
executed.  Thus, a hand-designed policy in a distributed systems can be replaced
with policy that has been optimized by a contextual learning system. By
supplying rich contexts to this policy, developers capture the complexity of our
environment without having to understand it. Additionally, by optimizing the
policy continuously, the learning system can swiftly adapt to changes in the
environment or service.

Machine learning has certainly been used in the past to optimize system
decisions. Most of this use has been restricted to either offline settings or
expect {\em full information}, whereby a supervised learner is trained on data
that has been annotated with correct labels.
Some systems work in an online setting with full information. An example of that
is a caching layer that uses a learning system to dynamically switch between
existing (hand-designed) cache eviction policies, by running all policies in
parallel and measuring their performance.
Systems that work in an online setting with partial information
tend to be for news or ad recommendation, but examples from systems
infrastructure exist. For example, caching dynamic query content [], tuning
parameters of multicore data structures [], allocating resources to cluster
computing tasks [], and allocating servers to web applications [].  In all of
these systems, the context used to make a decision, and the decision itself, are
{\em local} to a machine. This is achieved by either devising a decentralized
algorithm in the first place, or giving up on global optimality. We are not
aware of any system that makes ``global'' decisions based on truly distributed
architecture or state.

%% Hussam: I don't know what is this "and yet" referencing. So I am taking it out.
%%
%%And yet we use hand-designed policies to make these decisions all the time, such
%%as for request routing, replica placement, replica selection, cloud resource
%%allocation, failure recovery tasks, and others.

What is it about distributed systems that makes it difficult to apply online
learning at decision points such as request routing, replica placement, cloud
resource allocation, and failure recovery? This proposal identifies fundamental
challenges posed by distributed systems that must be analyzed and addressed in
order to make online learning a standard tool for distributed infrastructure
systems.  These challenges hold even if we assume that the online learning
system is a black-box that delivers optimal decisions with zero latency. In
other words, they are fundamental to the way we design distributed systems.  We
begin by highlighting some of the factors that contribute to these challenges,
and then list the challenges themselves.

@@

Systems often make decisions without complete knowledge of the factors that
affect, or could affect, those decisions.  For example, a caching system evicts
items based on what it believes future requests will look like, without actually
knowing the future.  In other cases, incomplete knowledge is due to the absence
of data-computation locality.  For example, a load balancer makes decisions
based on what it thinks is the current load on other machines, e.g. based on
data it collected recently, rather than the actual load at that very instant in
time.

In both of these situations, assumptions are made about the unknowns offline
during \emph{design time}, resulting in different request routing and caching
algorithms. A change to the \emph{run-time} environment could easily render
these choices suboptimal.

@@

In some situations, system designers have no way of knowing deployment details
that might impact a system's performance. For example, a system could be built
for external deployment and used by third-party customers, or it could be
deployed in a multi-tenant environment where co-located services could affect
its performance in unexpected ways. In these cases, it is nearly impossible for
developers to account for all unknown variables at design time.

@@

As mentioned earlier, we adopt the framework of contextual online learning with
partial feedback.  To separate the systems concerns we are trying to highlight
from machine learning concerns, we abstract the learning system as a black-box
and assume it provides optimal decisions with zero latency. For a given decision
over a discrete action space, the interface to the learning system is as
follows: given a {\em context} that summarizes the state of the world, the
learning system chooses an {\em action} to take, which our system does and later
reports a {\em reward} indicating how good the action was. An example of a
learning system that supports this interface is the Decision
Service~\footnote{http://aka.ms/mwt}. Now consider a load balancer using that
learning system to make its request-routing decisions. The context for each
decision could be load information of each machine in the cluster, the action
could be a specific machine to route the request to, and the reward could be the
latency of processing the request.

Thus we have reduced our concerns to generating an appropriate context,
executing an action (told to us by the learning system), and reporting reward
information.  We believe that fundamental properties of distributed systems make
even these tasks difficult, as described below.

@@

\subsection*{Hierarchy}
Distributed systems are often designed in a hierarchical manner, where
subcomponents are monitoring and managing others. These components and layers of
indirection simplify reasoning about the system at large. However, it is not
clear how hierarchy will affect an ML agent.
\begin{itemize}
  \item Will hierarchy aid the learning system via parallelism? Or will it hinder
    it since not all features are visible at every level?
  \item Should learning happen in each component of the system? Or should it be
    centralized?
  \item If learning is happening at multiple levels of the hierarchy, will that
    create interference between the learners? Or will they complement each
    other?
   \item Should feature vectors cut across the levels or not?
\end{itemize}

- In general should be composable if at each level can define decision and reward,
and exploration randomization is different at each level
